---
title: "Data analysis Income"
subtitle: "Changes in the Justification of Pension Inequality in Chile (2016–2023) and its Relationship to Social Class and Beliefs in Meritocracy"
author: "Andreas Laffert"
date: today
lang: en
fontsize: 12pt
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    toc: true
    toc-location: right
    toc-depth: 2
    toc-expand: 2
    toc-title: Contents
    number-sections: true
    number-depth: 3
    theme:
      - cosmo
      - edumer_html.scss
    code-link: true
    title-block-banner: true
editor_options: 
  chunk_output_type: console
---

# Presentation

This is the analysis code for the paper "Changes in the Justification of Pension Inequality in Chile (2016–2023) and its Relationship to Social Class and Beliefs in Meritocracy". The dataset used is `df_balanced_income.RData`.

```{r}
#| label: set
#| echo: false
#| message: false
#| warning: false

library(knitr)
knitr::opts_chunk$set(echo = TRUE, include = TRUE, warning = FALSE, message = FALSE)

table_format <- if(is_html_output()) {
  "html"
} else if(is_latex_output()) {
  "latex"
}
table_format2 <- if(is_html_output()) {
  T
} else if(is_latex_output()) {
  F
}

options(kableExtra.html.bsTable = T)
options(knitr.kable.NA = "")
```


# Libraries

```{r}
#| label: packages
#| include: true

if (! require("pacman")) install.packages("pacman")

pacman::p_load(tidyverse, 
               sjmisc, 
               sjPlot, 
               lme4, 
               here, 
               performance,
               influence.ME, 
               srvyr,
               ordinal,
               texreg, 
               ggdist,
               misty,
               kableExtra,
               ggalluvial, 
               shadowtext,
               MetBrewer,
               patchwork,
               sjlabelled)


options(scipen=999)
rm(list = ls())
```

# Data

```{r}
#| label: data
#| echo: true

load(file = here("input/data/rob_income/df_balanced_rob_income.RData"))

glimpse(df_balanced)

# Generate analytical sample

df_study1 <- df_balanced %>%
  select(-muestra) %>% 
  na.omit() %>% 
  mutate(ola = case_when(ola == "2016" ~ 1,
                         ola == "2017" ~ 2, 
                         ola == "2018" ~ 3,
                         ola == "2019" ~ 4,
                         ola == "2022" ~ 5,
                         ola == "2023" ~ 6)) %>% 
  mutate(ola = as.factor(ola),
         ola_num = as.numeric(ola),
         ola_2=as.numeric(ola)^2)

df_study1 <- df_study1 %>%
  group_by(idencuesta) %>%             # Agrupar por el identificador del participante
  mutate(n_participaciones = n()) %>%  # Contar el número de filas (participaciones) por participante
  ungroup()

df_study1 <- df_study1 %>% filter(n_participaciones>1)

# Corregir etiquetas

df_study1$just_pension <- sjlabelled::set_label(df_study1$just_pension, 
                        label = "Pension distributive justice")

df_study1$egp <- sjlabelled::set_label(df_study1$egp, 
                        label = "Social class")

df_study1$merit_effort <- sjlabelled::set_label(df_study1$merit_effort, 
                        label = "People are rewarded for their efforts")

df_study1$merit_talent <- sjlabelled::set_label(df_study1$merit_talent, 
                        label = "People are rewarded for their intelligence")


```

# Analysis

## Descriptives

```{r}
#| label: fig-alluvial
#| out-width: '100%'
#| fig-asp: 1
#| fig-cap: "Changes in the justification of inequality in pensions over time (2016-2023)"
#| fig-cap-location: top

datos.pension <- df_study1 %>% 
   mutate(just_pension = factor(just_pension, 
                              levels = c("Strongly agree",
                                         "Agree",
                                         "Neither agree nor disagree",
                                         "Disagree",
                                         "Strongly disagree"))) %>% 
  group_by(idencuesta, ola) %>% 
  count(just_pension) %>% 
  group_by(ola) %>% 
  mutate(porcentaje=n/sum(n)) %>% 
  ungroup() %>% 
  na.omit() %>% 
  mutate(wave = case_when(ola == 1 ~ "2016",
                          ola == 2 ~ "2017",
                          ola == 3 ~ "2018",
                          ola == 4 ~ "2019",
                          ola == 5 ~ "2022",
                          ola == 6 ~ "2023"),
         wave = factor(wave, levels = c("2016",
                                        "2017",
                                        "2018",
                                        "2019",
                                        "2022",
                                        "2023")))



etiquetas.pension <- df_study1 %>%
  mutate(just_pension = factor(just_pension, 
                              levels = c("Strongly agree",
                                         "Agree",
                                         "Neither agree nor disagree",
                                         "Disagree",
                                         "Strongly disagree"))) %>% 
  group_by(ola, just_pension) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(ola) %>%
  mutate(porcentaje = count / sum(count)) %>% 
  na.omit() %>% 
  mutate(idencuesta = 1,
         wave = case_when(ola == 1 ~ "2016",
                          ola == 2 ~ "2017",
                          ola == 3 ~ "2018",
                          ola == 4 ~ "2019",
                          ola == 5 ~ "2022",
                          ola == 6 ~ "2023"),
         wave = factor(wave, levels = c("2016",
                                        "2017",
                                        "2018",
                                        "2019",
                                        "2022",
                                        "2023")))


datos.pension %>% 
  ggplot(aes(x = wave, fill = just_pension, stratum = just_pension,
             alluvium = idencuesta, y = porcentaje)) +
  ggalluvial::geom_flow(alpha = .4) + 
  ggalluvial::geom_stratum(linetype = 0) +
  scale_y_continuous(labels = scales::percent) + 
  scale_fill_manual(values =  c("#0571B0","#92C5DE","#b3b3b3ff","#F4A582","#CA0020")) +
  geom_shadowtext(data = etiquetas.pension,
                  aes(label = ifelse(porcentaje > 0 , scales::percent(porcentaje, accuracy = .1),"")),
                  position = position_stack(vjust = .5),
                  show.legend = FALSE,
                  size = 3,
                  color = rep('white'),
                  bg.colour='grey30')+
  labs(y = "%",
       x = NULL,
       fill = NULL,
       title = NULL) +
  theme_ggdist() +
  theme(legend.position = "bottom") 

```


## Longitudinal multilevel models

```{r}
#| include: false

df_study1$merit_effort <- as_numeric(df_study1$merit_effort)
df_study1$merit_talent <- as_numeric(df_study1$merit_talent)

df_study1$decile_eq <- as_numeric(df_study1$decile_eq)
df_study1$educ <- if_else(df_study1$educ == "Universitary", 1, 0)

df_study1$merit_effort_dic <- if_else(df_study1$merit_effort <= 3, 0, 1)
df_study1$merit_talent_dic <- if_else(df_study1$merit_talent <= 3, 0, 1)

df_study1$age_2 <- (df_study1$age)^2 

df_study1 <- df_study1 %>% 
  mutate(ola = case_when(ola == 1 ~ "2016",
                         ola == 2 ~ "2017",
                         ola == 3 ~ "2018",
                         ola == 4 ~ "2019",
                         ola == 5 ~ "2022",
                         ola == 6 ~ "2023"),
         ola = factor(ola, levels = c("2016",
                                        "2017",
                                        "2018",
                                        "2019",
                                        "2022",
                                        "2023")))


df_study1 <- df_study1 %>% 
  group_by(idencuesta) %>% 
  mutate(merit_effort_mean = mean(merit_effort, na.rm = T),
         merit_effort_cwc = merit_effort - merit_effort_mean,
         merit_talent_mean = mean(merit_talent, na.rm = T),
         merit_talent_cwc = merit_talent - merit_talent_mean,
         
         educ_mean = mean(educ, na.rm = T),
         educ_cwc = educ - educ_mean,
         
         decile_eq_mean = mean(decile_eq, na.rm = T),
         decile_eq_cwc = decile_eq - decile_eq_mean,
         
         ) %>% 
  ungroup()

df_study1$just_pension <- as_factor(df_study1$just_pension)


```

## ICC 
```{r}
m0 <- clmm(just_pension ~ 1 + (1 | idencuesta), 
           link = "logit",
  Hess = TRUE, # calcula explícitamente la matriz varianza-covarianza de estimadores
                data = df_study1)

performance::icc(m0, by_group = T) # 0.23 es between, 0.77 within
```

## Time effects

```{r}
#| results: asis

#m1.1 <- clmm(just_pension ~ 1 + ola + (1 | idencuesta),
#                link = "logit",
#  Hess = TRUE,
#  data = df_study1)
#
#m1.2 <- clmm(just_pension ~ 1 + ola_num + (1 | idencuesta),
#                link = "logit",
#  Hess = TRUE, data = df_study1)
#
#m1.3 <- clmm(just_pension ~ 1 + ola_num + ola_2 + (1| idencuesta),
#                link = "logit",
#  Hess = TRUE, data = df_study1)
#
#m1.4 <- clmm(just_pension ~ 1 + ola_num + ola_2 + (1 + ola_num | #idencuesta),
#               link = "logit",
# Hess = TRUE, data = df_study1)
#
#save(m1.1,m1.2,m1.3,m1.4, file = here("output/rob_income/time_effects.RData"))

load(file = here("output/rob_income/time_effects.RData"))

ccoef <- list(
  "Strongly disagree|Disagree" = "Strongly disagree|Disagree",
  "Disagree|Neither agree nor disagree" = "Disagree|Neither agree nor disagree", 
  "Neither agree nor disagree|Agree" = "Neither agree nor disagree|Agree", 
  "Agree|Strongly agree" = "Agree|Strongly agree", 
  "ola2017" = "Wave 2017",
  "ola2018" = "Wave 2018",
  "ola2019" = "Wave 2019",
  "ola2022" = "Wave 2022",
  "ola2023" = "Wave 2023",
  ola_num = "Wave",
  ola_2 = "Wave^2")

texreg::htmlreg(list(m1.1,m1.2,m1.3,m1.4),
               caption.above = T,
               caption = NULL,
               stars = c(0.05, 0.01, 0.001),
               custom.coef.map = ccoef,
               digits = 3,
               groups = list("Wave (Ref.= 2016)" = 5:9),
               custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. %stars.",
               leading.zero = T,
               use.packages = F,
               booktabs = F,
               scalebox = 0.80,
               include.loglik = FALSE,
               include.aic = FALSE,
               center = T)
               
```

### Anova

```{r}
anova(m1.3, m1.4) # quedarse con tiempo continua y con pendiente aleatoria
```


## WE and BE main effects

```{r}
#| results: asis
## WE and BE main effects
#m3 <- clmm(just_pension ~ 1 + ola_num + ola_2 +   
#             merit_effort_cwc + merit_talent_cwc + (1 + ola_num | idencuesta), 
#            link = "logit",
#           Hess = TRUE,
#           data = df_study1)
#
#m4 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + 
#             merit_effort_cwc + merit_talent_cwc + 
#             merit_effort_mean + merit_talent_mean + 
#             (1 + ola_num| idencuesta), 
#            link = "logit",
#           Hess = TRUE,
#           data = df_study1)
#
#m5 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + 
#              merit_effort_cwc + merit_talent_cwc + 
#              merit_effort_mean + merit_talent_mean + 
#              decile_eq_mean +  decile_eq_cwc +
#              (1 + ola_num| idencuesta), 
#            link = "logit",
#            Hess = TRUE,
#           data = df_study1)
#
#m6 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + 
#              merit_effort_cwc + merit_talent_cwc + 
#              merit_effort_mean + merit_talent_mean + 
#              decile_eq_mean +  decile_eq_cwc +
#             educ_mean + educ_cwc +
#              (1 + ola_num| idencuesta), 
#            link = "logit",
#            Hess = TRUE,
#           data = df_study1)
#
#m7 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + 
#              merit_effort_cwc + merit_talent_cwc + 
#              merit_effort_mean + merit_talent_mean + 
#              decile_eq_mean +  decile_eq_cwc +
#             educ_mean + educ_cwc + ideo + sex + age + 
#              (1 + ola_num| idencuesta), 
#            link = "logit",
#            Hess = TRUE,
#           data = df_study1)
#

#save(m3,m4,m5,m6,m7, file = here("output/rob_income/main_effects.RData"))

load(file = here("output/rob_income/main_effects.RData"))

htmlreg(list(m3,m4,m5,m6,m7))

```


```{r echo=FALSE, results='asis'}
#| label: tbl-modelos
#| tbl-cap: "Cumulative link longitudinal multilevel models for pension market justice preferences"
#| tbl-cap-location: top

omit <- "(ideo)|(sex)|(age)"

ccoef <- list(
  #"Strongly disagree|Disagree" = "Strongly disagree|Disagree",
  #"Disagree|Neither agree nor disagree" = "Disagree|Neither agree nor disagree",
  #"Neither agree nor disagree|Agree" = "Neither agree nor disagree|Agree",
  #"Agree|Strongly agree" = "Agree|Strongly agree",
  "ola2017" = "Wave 2017",
  "ola2018" = "Wave 2018",
  "ola2019" = "Wave 2019",
  "ola2022" = "Wave 2022",
  "ola2023" = "Wave 2023",
  ola_num = "Wave",
  ola_2 = "Wave^2",
  merit_effort_cwc = "Merit: Effort (WE)",
  merit_talent_cwc = "Merit: Talent (WE)",
  merit_effort_mean = "Merit: Effort (BE)",
  merit_talent_mean = "Merit: Talent (BE)",
  decile_eq_mean = "Income (BE)",
  decile_eq_cwc = "Income (WE)",
  educ_mean = "Education (BE)",
  educ_cwc = "Education (WE)"
  )

htmlreg(list(m1.1, m1.4, m3, m4, m5, m6, m7),
               custom.model.names = c(paste0("Model ", seq(1:7))),
               caption.above = T,
               caption = NULL,
               stars = c(0.05, 0.01, 0.001),
               omit.coef = omit,
               custom.coef.map = ccoef,
               digits = 3,
               custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. %stars.",
               leading.zero = T,
               use.packages = F,
               booktabs = F,
               scalebox = 0.80,
               include.loglik = FALSE,
               include.aic = FALSE,
               center = T,
               custom.gof.names = c("BIC", "Numb. obs.", "Num. groups: individuals", "Var: individuals (Intercept)", "Var: individuals, wave"),
               custom.gof.rows = list("Controls"=c(rep("No",6), "Yes")))
               

```

```{r}
#| eval: false
#| include: false
sjPlot::plot_model(
  m7,
  type = "est",
  terms = c("ola_num",
            "ola_2",
            "merit_effort_cwc",
            "merit_talent_cwc", 
            "merit_effort_mean",
            "merit_talent_mean",
            "decile_eq_mean" ,
            "decile_eq_cwc",
            "educ_mean" ,
            "educ_cwc"),
  axis.labels = c( "Education (WE)",
                   "Education (BE)",
                   "Income (WE)",
                   "Income (BE)",
                   "Merit: Talent (BE)",
                  "Merit: Effort (BE)",
                  "Merit: Talent (WE)",
                  "Merit: Effort (WE)",
                  "Wave^2",
                  "Wave"),
  transform = NULL,
  show.ci = TRUE,     # <— muestra intervalos
  ci.lvl  = 0.95,     # <— nivel de confianza
  show.values = TRUE,
  p.style = "stars",
  vline.color = "grey40",
  value.size = 4.5
) +
  scale_y_continuous(limits = c(-1,1.5))+
  labs(title = NULL, 
       caption = "Source: own elaboration based on ELSOC data 2016-2023 (N obs = 6,123; N individuals = 1,327)\nControls: gender, age, and political identification\n*** p < 0.001; ** p < 0.01; * p < 0.05") +
  theme_ggdist() +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"),
        text = element_text(size = 16),
        plot.caption = element_text(size = 12, face = "italic"))

ggsave(
  plot = last_plot(),
  filename = here("presentations/in_equality-Conference-2026/images/income_regression.png"),
  device = "png",
  dpi = "retina",
  units = "cm",
  width = 33,
  height = 20
)

```




## Interactions with controls (direct effect)

```{r}
#| results: asis
 ## WE and BE Interactions with controls

#m11 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + 
#               merit_effort_cwc + merit_talent_cwc + 
#               merit_effort_mean + merit_talent_mean + 
#               decile_eq_mean*merit_effort_cwc + 
#             educ_mean + educ_cwc + ideo + sex + age + 
#               (1 + ola_num + merit_effort_cwc| idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
#m12 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + 
#               merit_effort_cwc + merit_talent_cwc + 
#               merit_effort_mean + merit_talent_mean + 
#               decile_eq_cwc*merit_effort_cwc + 
#               educ_mean + educ_cwc + ideo + sex + age +  (1 + #ola_num + merit_effort_cwc| idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
#
#m13 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + 
#               merit_effort_cwc + merit_talent_cwc + 
#               merit_effort_mean + merit_talent_mean + 
#               decile_eq_mean*merit_effort_mean + 
#             educ_mean + educ_cwc + ideo + sex + age + 
#               (1 + ola_num + merit_effort_mean| idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
#m14 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + 
#               merit_effort_cwc + merit_talent_cwc + 
#               merit_effort_mean + merit_talent_mean + 
#               decile_eq_cwc*merit_effort_mean + 
#               educ_mean + educ_cwc + ideo + sex + age +  (1 + #ola_num + merit_effort_mean| idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
#
#m15 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + 
#               merit_effort_cwc + merit_talent_cwc + 
#               merit_effort_mean + merit_talent_mean + 
#               decile_eq_mean*merit_talent_cwc + 
#               educ_mean + educ_cwc + ideo + sex + age + (1 + #ola_num + merit_talent_cwc| idencuesta), 
#           link = "logit",
# Hess = TRUE,
#          data = df_study1)
#
#m16 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + 
#               merit_effort_cwc + merit_talent_cwc + 
#               merit_effort_mean + merit_talent_mean + 
#               decile_eq_cwc*merit_talent_cwc + 
#               educ_mean + educ_cwc + ideo + sex + age +  (1 + #ola_num + merit_talent_cwc| idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
#m17 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + 
#               merit_effort_cwc + merit_talent_cwc + 
#               merit_effort_mean + merit_talent_mean + 
#               decile_eq_mean*merit_talent_mean + 
#               educ_mean + educ_cwc + ideo + sex + age + (1 + #ola_num + merit_talent_mean| idencuesta), 
#           link = "logit",
# Hess = TRUE,
#          data = df_study1)
#
#m18 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + 
#               merit_effort_cwc + merit_talent_cwc + 
#               merit_effort_mean + merit_talent_mean + 
#               decile_eq_cwc*merit_talent_mean + 
#               educ_mean + educ_cwc + ideo + sex + age +  (1 + #ola_num + merit_talent_mean| idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
#save(m11,m12,m13,m14,
#     m15, m16, m17,m18,file = #here("output/rob_income/interactions_direct.RData"))

load(file = here("output/rob_income/interactions_direct.RData"))

htmlreg(list(m11,m12,m13,m14))

htmlreg(list(m15,m16,m17,m18))
```

