---
title: "Data analysis 2"
subtitle: "Changes in the Justification of Pension Inequality in Chile (2016–2023) and its Relationship to Social Class and Beliefs in Meritocracy"
author: "Andreas Laffert & René Canales"
date: today
lang: en
fontsize: 12pt
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    toc: true
    toc-location: right
    toc-depth: 2
    toc-expand: 2
    toc-title: Contents
    number-sections: true
    number-depth: 3
    theme:
      - cosmo
      - edumer_html.scss
    code-link: true
    title-block-banner: true
editor_options: 
  chunk_output_type: console
---

# Presentation

This is the analysis code for the paper "Changes in the Justification of Pension Inequality in Chile (2016–2023) and its Relationship to Social Class and Beliefs in Meritocracy". The dataset used is `df_balanced.RData`.

```{r}
#| label: set
#| echo: false
#| message: false
#| warning: false

library(knitr)
knitr::opts_chunk$set(echo = TRUE, include = TRUE, warning = FALSE, message = FALSE)

table_format <- if(is_html_output()) {
  "html"
} else if(is_latex_output()) {
  "latex"
}
table_format2 <- if(is_html_output()) {
  T
} else if(is_latex_output()) {
  F
}

options(kableExtra.html.bsTable = T)
options(knitr.kable.NA = "")
```


# Libraries

```{r}
#| label: packages
#| include: true

if (! require("pacman")) install.packages("pacman")

pacman::p_load(tidyverse, 
               sjmisc, 
               sjPlot, 
               lme4, 
               here, 
               performance,
               influence.ME, 
               srvyr,
               ordinal,
               texreg, 
               ggdist,
               misty,
               kableExtra,
               ggalluvial, 
               shadowtext,
               MetBrewer,
               patchwork,
               sjlabelled)


options(scipen=999)
rm(list = ls())
```

# Data

```{r}
#| label: data
#| echo: true

load(file = here("input/data/proc/df_balanced.RData"))

glimpse(df_balanced)

# Generate analytical sample

df_study1 <- df_balanced %>%
  select(-muestra) %>% 
  na.omit() %>% 
  mutate(ola = case_when(ola == 1 ~ 1,
                         ola == 2 ~ 2, 
                         ola == 3 ~ 3,
                         ola == 4 ~ 4,
                         ola == 6 ~ 5,
                         ola == 7 ~ 6)) %>% 
  mutate(ola = as.factor(ola),
         ola_num = as.numeric(ola),
         ola_2=as.numeric(ola)^2)

df_study1 <- df_study1 %>%
  group_by(idencuesta) %>%             # Agrupar por el identificador del participante
  mutate(n_participaciones = n()) %>%  # Contar el número de filas (participaciones) por participante
  ungroup()

df_study1 <- df_study1 %>% filter(n_participaciones>1)

# Corregir etiquetas

df_study1$just_pension <- sjlabelled::set_label(df_study1$just_pension, 
                        label = "Pension distributive justice")

df_study1$egp <- sjlabelled::set_label(df_study1$egp, 
                        label = "Social class")

df_study1$merit_effort <- sjlabelled::set_label(df_study1$merit_effort, 
                        label = "People are rewarded for their efforts")

df_study1$merit_talent <- sjlabelled::set_label(df_study1$merit_talent, 
                        label = "People are rewarded for their intelligence")


```

# Analysis

## Descriptives

```{r}
#| label: fig-alluvial
#| out-width: '100%'
#| fig-asp: 1
#| fig-cap: "Changes in the justification of inequality in pensions over time (2016-2023)"
#| fig-cap-location: top

datos.pension <- df_study1 %>% 
   mutate(just_pension = factor(just_pension, 
                              levels = c("Strongly agree",
                                         "Agree",
                                         "Neither agree nor disagree",
                                         "Disagree",
                                         "Strongly disagree"))) %>% 
  group_by(idencuesta, ola) %>% 
  count(just_pension) %>% 
  group_by(ola) %>% 
  mutate(porcentaje=n/sum(n)) %>% 
  ungroup() %>% 
  na.omit() %>% 
  mutate(wave = case_when(ola == 1 ~ "2016",
                          ola == 2 ~ "2017",
                          ola == 3 ~ "2018",
                          ola == 4 ~ "2019",
                          ola == 5 ~ "2022",
                          ola == 6 ~ "2023"),
         wave = factor(wave, levels = c("2016",
                                        "2017",
                                        "2018",
                                        "2019",
                                        "2022",
                                        "2023")))



etiquetas.pension <- df_study1 %>%
  mutate(just_pension = factor(just_pension, 
                              levels = c("Strongly agree",
                                         "Agree",
                                         "Neither agree nor disagree",
                                         "Disagree",
                                         "Strongly disagree"))) %>% 
  group_by(ola, just_pension) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(ola) %>%
  mutate(porcentaje = count / sum(count)) %>% 
  na.omit() %>% 
  mutate(idencuesta = 1,
         wave = case_when(ola == 1 ~ "2016",
                          ola == 2 ~ "2017",
                          ola == 3 ~ "2018",
                          ola == 4 ~ "2019",
                          ola == 5 ~ "2022",
                          ola == 6 ~ "2023"),
         wave = factor(wave, levels = c("2016",
                                        "2017",
                                        "2018",
                                        "2019",
                                        "2022",
                                        "2023")))


datos.pension %>% 
  ggplot(aes(x = wave, fill = just_pension, stratum = just_pension,
             alluvium = idencuesta, y = porcentaje)) +
  ggalluvial::geom_flow(alpha = .4) + 
  ggalluvial::geom_stratum(linetype = 0) +
  scale_y_continuous(labels = scales::percent) + 
  scale_fill_manual(values =  c("#0571B0","#92C5DE","#b3b3b3ff","#F4A582","#CA0020")) +
  geom_shadowtext(data = etiquetas.pension,
                  aes(label = ifelse(porcentaje > 0 , scales::percent(porcentaje, accuracy = .1),"")),
                  position = position_stack(vjust = .5),
                  show.legend = FALSE,
                  size = 3,
                  color = rep('white'),
                  bg.colour='grey30')+
  labs(y = "%",
       x = NULL,
       fill = NULL,
       title = NULL) +
  theme_ggdist() +
  theme(legend.position = "bottom") 

```


```{r}
#| label: tbl-summary2
#| tbl-cap: "Estadísticos descriptivos variables independientes"
#| tbl-cap-location: top
t2 <- df_study1 %>% 
  filter(ola == last(ola)) %>% 
  select(just_pension, egp, merit_effort, merit_talent) 

print(summarytools::dfSummary(t2), method="render")
```


```{r}
#| label: fig-meanchange
#| out-width: '100%'
#| fig-asp: 1
#| fig-cap: "Changes in the standarized mean of market justice preferences in pensions and meritocracy by social class (2016-2023)"
#| fig-cap-location: top
#| echo: false
#| results: asis

df_pond <- df_study1 %>% 
  mutate(ola = case_when(ola == 1 ~ "2016",
                         ola == 2 ~ "2017",
                         ola == 3 ~ "2018",
                         ola == 4 ~ "2019",
                         ola == 5 ~ "2022",
                         ola == 6 ~ "2023"),
         ola = factor(ola, levels = c("2016",
                                      "2017",
                                      "2018",
                                      "2019",
                                      "2022",
                                      "2023"))) %>% 
  as_survey_design(.data = .,
                   ids = segmento, 
                   strata = estrato, 
                   weights = ponderador_long_total)

df_pond <- df_pond %>% 
  select(ola, egp, just_pension,  merit_effort, merit_talent) %>% 
  mutate_at(.vars = 3:5, .funs = ~ as.numeric(.)) %>% 
  mutate(
    just_pension_z    = as.numeric(scale(just_pension)),
    merit_effort_z    = as.numeric(scale(merit_effort)),
    merit_talent_z    = as.numeric(scale(merit_talent))
  )


df_pond %>% 
  select(ola, egp, just_pension_z,  merit_effort_z, merit_talent_z) %>% 
  group_by(ola, egp) %>% 
  summarise_all(~survey_mean(., vartype = "ci")) %>% 
  pivot_longer(cols = -c(ola, egp),
               names_to = "temp",
               values_to = "valor") %>%
  mutate(
    ci = case_when(
      str_ends(temp, "_low") ~ "ic_low",
      str_ends(temp, "_upp") ~ "ic_upp",
      TRUE                   ~ "mean"
    ),
    variable = str_remove(temp, "_low|_upp")
  ) %>%
  select(ola, egp, variable, ci, valor) %>%
  pivot_wider(
    names_from  = ci,
    values_from = valor
  ) %>% 
  mutate(variable = case_when(variable == "just_pension_z" ~ "Justification of pension inequality",
                              variable == "merit_effort_z"     ~ "Merit: Effort",
                              variable == "merit_talent_z"     ~ "Merit: Talent"),
         variable = factor(variable, levels = c("Justification of pension inequality",
                                                "Merit: Effort",
                                                "Merit: Talent")
         )) %>% 
  ggplot(aes(x = ola, y = mean, group = egp)) +
  geom_point(aes(shape=egp, color=egp), size = 3.5) +
  geom_line(aes(color = egp), linewidth = 0.8) +
  geom_errorbar(aes(ymin = ic_low, ymax = ic_upp, color = egp),
                width = 0.1) +
  scale_y_continuous(limits = c(-1, 1), n.breaks = 5) +
  facet_wrap(~variable, nrow = 3) +
  labs(y = "Mean (Z-score)",
       x = "Wave",
       color = NULL,
       shape = NULL,
       caption = "Source: own elaboration with pooled data from ELSOC 2016-2023 (N obs = 7,522; N groups = 1,317)") +
  theme_ggdist() +
  theme(legend.position = "top",
        text = element_text(size = 12)) 
 
```


## Longitudinal multilevel models

```{r}
#| include: false

df_study1$merit_effort <- as_numeric(df_study1$merit_effort)
df_study1$merit_talent <- as_numeric(df_study1$merit_talent)

df_study1$merit_effort_dic <- if_else(df_study1$merit_effort <= 3, 0, 1)
df_study1$merit_talent_dic <- if_else(df_study1$merit_talent <= 3, 0, 1)

df_study1$age_2 <- (df_study1$age)^2 

df_study1 <- df_study1 %>% 
  mutate(ola = case_when(ola == 1 ~ "2016",
                         ola == 2 ~ "2017",
                         ola == 3 ~ "2018",
                         ola == 4 ~ "2019",
                         ola == 5 ~ "2022",
                         ola == 6 ~ "2023"),
         ola = factor(ola, levels = c("2016",
                                        "2017",
                                        "2018",
                                        "2019",
                                        "2022",
                                        "2023")))


df_study1 <- df_study1 %>% 
  group_by(idencuesta) %>% 
  mutate(merit_effort_mean = mean(merit_effort, na.rm = T),
         merit_effort_cwc = merit_effort - merit_effort_mean,
         merit_talent_mean = mean(merit_talent, na.rm = T),
         merit_talent_cwc = merit_talent - merit_talent_mean,
         merit_effort_dic_mean = mean(merit_effort_dic, na.rm = T),
         merit_effort_dic_cwc = merit_effort_dic - merit_effort_dic_mean,
         merit_talent_dic_mean = mean(merit_talent_dic, na.rm = T),
         merit_talent_dic_cwc = merit_talent_dic - merit_talent_dic_mean,
         ) %>% 
  ungroup()

df_study1$just_pension <- as_factor(df_study1$just_pension)

df_study1$egp <- factor(df_study1$egp, levels = c("Working class (V+VI+VII)",
                                                  "Intermediate class (III+IV)",
                                                  "Service class (I+II)",
                                                  "Retired or pensioner", 
                                                  "Unemployed",
                                                  "Performs unpaid tasks"))
```

## ICC 
```{r}
m0 <- clmm(just_pension ~ 1 + (1 | idencuesta), 
           link = "logit",
  Hess = TRUE, # calcula explícitamente la matriz varianza-covarianza de estimadores
                data = df_study1)

performance::icc(m0, by_group = T) # 0.24 es between, 0.76 within
```

## Time effects

```{r}
#| results: asis

#m1.1 <- clmm(just_pension ~ 1 + ola + (1 | idencuesta),
#                link = "logit",
#  Hess = TRUE,
#  data = df_study1)


#m1.2 <- clmm(just_pension ~ 1 + ola_num + (1 | idencuesta),
#                link = "logit",
#  Hess = TRUE, data = df_study1)
#
#m1.3 <- clmm(just_pension ~ 1 + ola_num + ola_2 + (1| idencuesta),
#                link = "logit",
#  Hess = TRUE, data = df_study1)
#
#m1.4 <- clmm(just_pension ~ 1 + ola_num + ola_2 + (1 + ola_num | idencuesta),
#                link = "logit",
#  Hess = TRUE, data = df_study1)
#
#save(m1.1,m1.2,m1.3,m1.4, file = here("output/time_effects.RData"))
#
load(file = here("output/time_effects.RData"))

ccoef <- list(
  "Strongly disagree|Disagree" = "Strongly disagree|Disagree",
  "Disagree|Neither agree nor disagree" = "Disagree|Neither agree nor disagree", 
  "Neither agree nor disagree|Agree" = "Neither agree nor disagree|Agree", 
  "Agree|Strongly agree" = "Agree|Strongly agree", 
  "ola2017" = "Wave 2017",
  "ola2018" = "Wave 2018",
  "ola2019" = "Wave 2019",
  "ola2022" = "Wave 2022",
  "ola2023" = "Wave 2023",
  ola_num = "Wave",
  ola_2 = "Wave^2")

texreg::screenreg(list(m1.1,m1.2,m1.3,m1.4),
               caption.above = T,
               caption = NULL,
               stars = c(0.05, 0.01, 0.001),
               custom.coef.map = ccoef,
               digits = 3,
               groups = list("Wave (Ref.= 2016)" = 5:9),
               custom.note = "Note: Cells contain regression coefficients with standard errors in parentheses. %stars.",
               leading.zero = T,
               use.packages = F,
               booktabs = F,
               scalebox = 0.80,
               include.loglik = FALSE,
               include.aic = FALSE,
               center = T)
               
```

### Anova

```{r}
anova(m1.3, m1.4) # quedarse con tiempo continua y con pendiente aleatoria
```


## WE and BE main effects

```{r}
#| results: asis
## WE and BE main effects
#m3 <- clmm(just_pension ~ 1 + ola_num + ola_2 + egp + 
#             (1 + ola_num | idencuesta), 
#           link = "logit",
#           Hess = TRUE,
#           data = df_study1)
#
#m4 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + egp +  
#             merit_effort_cwc + merit_talent_cwc + (1 + ola_num | idencuesta), 
#            link = "logit",
#           Hess = TRUE,
#           data = df_study1)
#
#m5 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + egp +  
#             merit_effort_cwc + merit_talent_cwc + 
#             merit_effort_mean + merit_talent_mean + 
#             (1 + ola_num| idencuesta), 
#            link = "logit",
#           Hess = TRUE,
#           data = df_study1)
#
#m6 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + egp + 
#              merit_effort_cwc + merit_talent_cwc + 
#              merit_effort_mean + merit_talent_mean + 
#              educ + ideo + sex + age + 
#              (1 + ola_num| idencuesta), 
#            link = "logit",
#            Hess = TRUE,
#           data = df_study1)
#
#save(m3,m4,m5,m6, file = here("output/main_effects.RData"))

load(file = here("output/main_effects.RData"))

screenreg(list(m3,m4,m5,m6))

```


```{r}
#| results: asis
## WE and BE main effects
#m3_oesch <- clmm(just_pension ~ 1 + ola_num + ola_2 + oesch + 
#             (1 + ola_num | idencuesta), 
#           link = "logit",
#           Hess = TRUE,
#           data = df_study1)
#
#m4_oesch <- clmm(just_pension ~ 1 + ola_num +  ola_2 + oesch +  
#             merit_effort_cwc + merit_talent_cwc + (1 + ola_num | idencuesta), 
#            link = "logit",
#           Hess = TRUE,
#           data = df_study1)
#
#m5_oesch <- clmm(just_pension ~ 1 + ola_num +  ola_2 + oesch +  
#             merit_effort_cwc + merit_talent_cwc + 
#             merit_effort_mean + merit_talent_mean + 
#             (1 + ola_num| idencuesta), 
#            link = "logit",
#           Hess = TRUE,
#           data = df_study1)
#
#m6_oesch <- clmm(just_pension ~ 1 + ola_num +  ola_2 + oesch + 
#              merit_effort_cwc + merit_talent_cwc + 
#              merit_effort_mean + merit_talent_mean + 
#              educ + ideo + sex + age + 
#              (1 + ola_num| idencuesta), 
#            link = "logit",
#            Hess = TRUE,
#           data = df_study1)
#
#save(m3_oesch,m4_oesch,m5_oesch,m6_oesch, file = here("output/oesch_main_effects.RData"))
#
load(file = here("output/oesch_main_effects.RData"))

htmlreg(list(m3_oesch,m4_oesch,m5_oesch,m6_oesch))

```


## Interactions without controls (total effect)

```{r}
#| results: asis

## WE and BE Interactions without controls

#m7 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + egp + 
#             merit_effort_cwc + merit_talent_cwc + 
#             merit_effort_mean + merit_talent_mean + 
#             egp*merit_effort_cwc + (1 + ola_num + merit_effort_cwc| idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
#m8 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + egp + 
#             merit_effort_cwc + merit_talent_cwc + 
#             merit_effort_mean + merit_talent_mean + 
#             egp*merit_talent_cwc + (1 + ola_num + merit_talent_cwc| idencuesta), 
#           link = "logit",
# Hess = TRUE,
#          data = df_study1)
#
#
#m9 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + egp + 
#             merit_effort_cwc + merit_talent_cwc + 
#             merit_effort_mean + merit_talent_mean + 
#             egp*merit_effort_mean + (1 + ola_num + merit_effort_mean| idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
#m10 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + egp + 
#              merit_effort_cwc + merit_talent_cwc + 
#              merit_effort_mean + merit_talent_mean +
#              egp*merit_talent_mean + (1 + ola_num + merit_talent_mean| idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
#save(m7,m8,m9,m10, file = here("output/interactions_total.RData"))
load(file = here("output/interactions_total.RData"))

htmlreg(list(m7, m8, m9, m10))


```


```{r}
#| results: asis
 ## WE and BE Interactions without controls

#m7_oesch <- clmm(just_pension ~ 1 + ola_num +  ola_2 + oesch + 
#             merit_effort_cwc + merit_talent_cwc + 
#             merit_effort_mean + merit_talent_mean + 
#             oesch*merit_effort_cwc + (1 + ola_num + merit_effort_cwc| idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
#m8_oesch <- clmm(just_pension ~ 1 + ola_num +  ola_2 + oesch + 
#             merit_effort_cwc + merit_talent_cwc + 
#             merit_effort_mean + merit_talent_mean + 
#             oesch*merit_talent_cwc + (1 + ola_num + merit_talent_cwc| idencuesta), 
#           link = "logit",
# Hess = TRUE,
#          data = df_study1)
#
#
#m9_oesch <- clmm(just_pension ~ 1 + ola_num +  ola_2 + oesch + 
#             merit_effort_cwc + merit_talent_cwc + 
#             merit_effort_mean + merit_talent_mean + 
#             oesch*merit_effort_mean + (1 + ola_num + merit_effort_mean| idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
#m10_oesch <- clmm(just_pension ~ 1 + ola_num +  ola_2 + oesch + 
#              merit_effort_cwc + merit_talent_cwc + 
#              merit_effort_mean + merit_talent_mean +
#              oesch*merit_talent_mean + (1 + ola_num + merit_talent_mean| idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
#save(m7_oesch,m8_oesch,m9_oesch,m10_oesch, file = here("output/oesch_interactions_total.RData"))

load(file = here("output/oesch_interactions_total.RData"))

htmlreg(list(m7_oesch, m8_oesch, m9_oesch, m10_oesch))


```




## Interactions with controls (direct effect)

```{r}
#| results: asis
 ## WE and BE Interactions with controls

# m11 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + egp + 
#               merit_effort_cwc + merit_talent_cwc + 
#               merit_effort_mean + merit_talent_mean + 
#               ideo + sex + age + 
#               egp*merit_effort_cwc + (1 + ola_num + merit_effort_cwc| idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
# m12 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + egp + 
#               merit_effort_cwc + merit_talent_cwc +
#               merit_effort_mean + merit_talent_mean + 
#               ideo + sex + age + 
#               egp*merit_talent_cwc + (1 + ola_num + merit_talent_cwc| idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
#
#m13 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + egp + 
#              merit_effort_cwc + merit_talent_cwc + 
#              merit_effort_mean + merit_talent_mean + 
#              ideo + sex + age + 
#              egp*merit_effort_mean + (1 + ola_num + merit_effort_mean| #idencuesta), 
#           link = "logit",
# Hess = TRUE,
#          data = df_study1)

#m14 <- clmm(just_pension ~ 1 + ola_num +  ola_2 + egp + 
#              merit_effort_cwc + merit_talent_cwc + 
#              merit_effort_mean + merit_talent_mean + 
#              ideo + sex + age + 
#              egp*merit_talent_mean + (1 + ola_num + merit_talent_mean| #idencuesta), 
#           link = "logit",
# Hess = TRUE,
#          data = df_study1)

#save(m11,m12,m13,m14, file = here("output/interactions_direct.RData"))

load(file = here("output/interactions_direct.RData"))

htmlreg(list(m11,m12,m13,m14))

```


```{r}
#| results: asis
 ## WE and BE Interactions with controls

# m11_oesch <- clmm(just_pension ~ 1 + ola_num +  ola_2 + oesch + 
#               merit_effort_cwc + merit_talent_cwc + 
#               merit_effort_mean + merit_talent_mean + 
#               ideo + sex + age + 
#               oesch*merit_effort_cwc + (1 + ola_num + merit_effort_cwc| idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
# m12_oesch <- clmm(just_pension ~ 1 + ola_num +  ola_2 + oesch + 
#               merit_effort_cwc + merit_talent_cwc +
#               merit_effort_mean + merit_talent_mean + 
#               ideo + sex + age + 
#               oesch*merit_talent_cwc + (1 + ola_num + merit_talent_cwc| idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
#
# m13_oesch <- clmm(just_pension ~ 1 + ola_num +  ola_2 + oesch + 
#               merit_effort_cwc + merit_talent_cwc + 
#               merit_effort_mean + merit_talent_mean + 
#               ideo + sex + age + 
#               oesch*merit_effort_mean + (1 + ola_num | idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
# m14_oesch <- clmm(just_pension ~ 1 + ola_num +  ola_2 + oesch + 
#               merit_effort_cwc + merit_talent_cwc + 
#               merit_effort_mean + merit_talent_mean + 
#               ideo + sex + age + 
#               oesch*merit_talent_mean + (1 + ola_num | idencuesta), 
#            link = "logit",
#  Hess = TRUE,
#           data = df_study1)
#
#save(m11_oesch,m12_oesch,m13_oesch,m14_oesch, file = #here("output/oesch_interactions_direct.RData"))

load(file = here("output/oesch_interactions_direct.RData"))

htmlreg(list(m11_oesch,m12_oesch,m13_oesch,m14_oesch))

```

### Predicted probabilities

```{r}
library(ggeffects)

p1 <- ggpredict(
  m12,
  terms = c(
    "merit_talent_cwc [all]",
    "egp [Working class (V+VI+VII), Unemployed]"
  ),
  bias_correction = T,
  type = "fixed"
)

p2 <- ggpredict(
  m14,
  terms = c(
    "merit_talent_mean [all]",
    "egp [Working class (V+VI+VII), Intermediate class (III+IV)]"
  ),
  bias_correction = T,
  type = "fixed"
)

save(p1, p2, file = here("output/predicted_probabilities.RData"))
```

