---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: packages
#| include: false
#| echo: false

if (! require("pacman")) install.packages("pacman")

pacman::p_load(tidyverse, 
               sjmisc, 
               sjPlot, 
               lme4, 
               here, 
               performance,
               influence.ME, 
               srvyr,
               ordinal,
               texreg, 
               ggdist,
               misty,
               kableExtra,
               ggalluvial, 
               shadowtext,
               MetBrewer,
               patchwork,
               sjlabelled)


options(scipen=999)
rm(list = ls())
```

```{r}
#| label: data
#| include: false
#| echo: false

load(file = here("input/data/proc/df_balanced.RData"))

glimpse(df_balanced)

# Generate analytical sample

df_study1 <- df_balanced %>%
  select(-muestra) %>% 
  na.omit() %>% 
  mutate(ola = case_when(ola == 1 ~ 1,
                         ola == 2 ~ 2, 
                         ola == 3 ~ 3,
                         ola == 4 ~ 4,
                         ola == 6 ~ 5,
                         ola == 7 ~ 6)) %>% 
  mutate(ola = as.factor(ola),
         ola_num = as.numeric(ola),
         ola_2=as.numeric(ola)^2)

df_study1 <- df_study1 %>%
  group_by(idencuesta) %>%             # Agrupar por el identificador del participante
  mutate(n_participaciones = n()) %>%  # Contar el número de filas (participaciones) por participante
  ungroup()

df_study1 <- df_study1 %>% filter(n_participaciones>1)

# Corregir etiquetas

df_study1$just_pension <- sjlabelled::set_label(df_study1$just_pension, 
                        label = "Pension distributive justice")

df_study1$egp <- sjlabelled::set_label(df_study1$egp, 
                        label = "Social class")

df_study1$merit_effort <- sjlabelled::set_label(df_study1$merit_effort, 
                        label = "People are rewarded for their efforts")

df_study1$merit_talent <- sjlabelled::set_label(df_study1$merit_talent, 
                        label = "People are rewarded for their intelligence")


```

# Methodology {background-color="#491106ff"}

## Data

::: {.incremental .highlight-last style="font-size: 120%;"}

- [ELSOC (COES)](https://coes.cl/elsoc/): a representative panel survey of the urban adult population in Chile, based on a probabilistic, stratified, clustered, and multistage sampling design across large and small cities

- Period: six waves (2016, 2017, 2018, 2019, 2022, and 2023)

- Attrition 2016→2023: ~ 40%

- Analytical sample (**balanced 6-wave panel**): 

    * **N = 7,522 observations** (level 1) nested within 
    * **N = 1,317 individuals** (level 2)

:::




## Dependent variable
```{r}
#| label: fig-desc
#| fig-cap-location: bottom
#| fig-cap: "Distribution of responses for preferences for market justice in pensions in Chile (2016–2023)"
#| fig-align: center
#| out-width: '90%'
#| fig-asp: 0.6
#| echo: false

## Alluvial
datos.pension <- df_study1 %>% 
  mutate(just_pension = factor(just_pension, 
                               levels = c("Strongly agree",
                                          "Agree",
                                          "Neither agree nor disagree",
                                          "Disagree",
                                          "Strongly disagree"))) %>% 
  group_by(idencuesta, ola) %>% 
  count(just_pension) %>% 
  group_by(ola) %>% 
  mutate(porcentaje=n/sum(n)) %>% 
  ungroup() %>% 
  na.omit() %>% 
  mutate(wave = case_when(ola == 1 ~ "2016",
                          ola == 2 ~ "2017",
                          ola == 3 ~ "2018",
                          ola == 4 ~ "2019",
                          ola == 5 ~ "2022",
                          ola == 6 ~ "2023"),
         wave = factor(wave, levels = c("2016",
                                        "2017",
                                        "2018",
                                        "2019",
                                        "2022",
                                        "2023")))



etiquetas.pension <- df_study1 %>%
  mutate(just_pension = factor(just_pension, 
                               levels = c("Strongly agree",
                                          "Agree",
                                          "Neither agree nor disagree",
                                          "Disagree",
                                          "Strongly disagree"))) %>% 
  group_by(ola, just_pension) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(ola) %>%
  mutate(porcentaje = count / sum(count)) %>% 
  na.omit() %>% 
  mutate(idencuesta = 1,
         wave = case_when(ola == 1 ~ "2016",
                          ola == 2 ~ "2017",
                          ola == 3 ~ "2018",
                          ola == 4 ~ "2019",
                          ola == 5 ~ "2022",
                          ola == 6 ~ "2023"),
         wave = factor(wave, levels = c("2016",
                                        "2017",
                                        "2018",
                                        "2019",
                                        "2022",
                                        "2023")))


datos.pension %>% 
  ggplot(aes(x = wave, fill = just_pension, stratum = just_pension,
             alluvium = idencuesta, y = porcentaje)) +
  ggalluvial::geom_flow(alpha = .4) + 
  ggalluvial::geom_stratum(linetype = 0) +
  scale_y_continuous(labels = scales::percent) + 
  scale_fill_manual(values =  c("#0571B0","#92C5DE","#b3b3b3ff","#F4A582","#CA0020")) +
  geom_text(data = etiquetas.pension,
            aes(label = ifelse(porcentaje > 0 , scales::percent(porcentaje, accuracy = .1),"")),
            position = position_stack(vjust = .5),
            show.legend = FALSE, fontface = "bold",
            size = 4,
            color = 'white') +
  labs(y = NULL,
       x = NULL,
       fill = NULL,
       title = "It is fair that people with higher incomes have better pensions than people\nwith lower incomes?",
       caption = "Source: own elaboration with data from ELSOC 2016-2023 (N obs = 7,522; N individuals = 1,317)") +
  theme_ggdist() +
  theme(legend.position = "top",
        plot.title = element_text(face = "bold"),
        text = element_text(size = 16)) 

```


## Independent variables: Level 1

```{r}
#| label: fig-merit
#| fig-cap-location: bottom
#| fig-cap: "Distribution of responses for meritocracy items"
#| fig-align: center
#| out-width: '90%'
#| fig-asp: 0.6
#| echo: false


library(scales)
library(forcats)

df_plot <- df_study1 %>%
  select(merit_effort, merit_talent) %>%
  pivot_longer(everything(), names_to = "item", values_to = "x") %>%
  mutate(
    item = factor(
      item,
      levels = c("merit_effort", "merit_talent"),
      labels = c(
        "People are rewarded for their efforts",
        "People are rewarded for their \nintelligence and ability"
      )
    ),
    # x ya viene etiquetada; solo la normalizamos como character -> factor
    cat = as.character(x),
    cat = factor(
      cat,
      levels = c(
        "Strongly disagree",
        "Disagree",
        "Neither agree nor disagree",
        "Agree",
        "Strongly agree"
      )
    )
  ) %>%
  drop_na(cat) %>%
  count(item, cat) %>%
  group_by(item) %>%
  mutate(p = n / sum(n)) %>%
  ungroup()

wide <- df_plot %>%
  select(item, cat, p) %>%
  pivot_wider(names_from = cat, values_from = p, values_fill = 0)

# neutral separado a la izquierda
neutral_rect <- wide %>%
  transmute(
    item,
    cat = "Neither agree nor disagree",
    xmin = -1.08,
    xmax = -1.08 + `Neither agree nor disagree`,
    xmid = -1.08 + (`Neither agree nor disagree` / 2),
    p = `Neither agree nor disagree`
  )

# bloque principal divergente
main_rect <- bind_rows(
  wide %>%
    transmute(
      item,
      cat = "Strongly agree",
      xmin = -(`Agree` + `Strongly agree`),
      xmax = -`Agree`,
      p = `Strongly agree`
    ),
  wide %>%
    transmute(
      item,
      cat = "Agree",
      xmin = -`Agree`,
      xmax = 0,
      p = `Agree`
    ),
  wide %>%
    transmute(
      item,
      cat = "Disagree",
      xmin = 0,
      xmax = `Disagree`,
      p = `Disagree`
    ),
  wide %>%
    transmute(
      item,
      cat = "Strongly disagree",
      xmin = `Disagree`,
      xmax = `Disagree` + `Strongly disagree`,
      p = `Strongly disagree`
    )
) %>%
  mutate(xmid = (xmin + xmax) / 2)

rects <- bind_rows(main_rect, neutral_rect) %>%
  mutate(
    cat = factor(
      cat,
      levels = c(
        "Strongly agree",
        "Agree",
        "Neither agree nor disagree",
        "Disagree",
        "Strongly disagree"
      )
    ),
    label = percent(p, accuracy = 0.1))


rects[1,3] <- (-0.29)
rects[2,3] <- (-0.35)

ggplot(rects) +
  geom_rect(
    aes(
      xmin = xmin, xmax = xmax,
      ymin = as.numeric(item) - 0.25,
      ymax = as.numeric(item) + 0.25,
      fill = cat
    ),
    linewidth = 0.5
  ) +
  geom_vline(xintercept = 0, color = "gray40", linewidth = 0.4) +
  geom_text(
    aes(x = xmid, y = item, label = label),
    color = "white", fontface = "bold",
    size = 4
  ) +
  scale_fill_manual(values = c(
    "Strongly disagree" = "#CA0020",
    "Disagree" = "#F4A582",
    "Agree" = "#92C5DE",
    "Strongly agree" =  "#0571B0",
    "Neither agree nor disagree" = "gray65"
  )) +
  scale_x_continuous(
    limits = c(-1.15, 1),
    breaks = seq(-1, 1, by = 0.2),
    labels = function(x) percent(abs(x), accuracy = 1)
  ) +
  labs(
    x = NULL,
    y = NULL,
    fill = NULL,
    caption = "Source: own elaboration with data from ELSOC 2016-2023 (N obs = 7,522; N individuals = 1,317)"
  ) +
  theme_ggdist(base_size = 16) +
  theme(
    legend.position = "top",
    axis.text.y = element_text(face = "bold"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank()
  )

```




## Independent variables: Level 2


```{r}
#| label: fig-class
#| fig-cap-location: bottom
#| fig-cap: "Frecuency of class position"
#| fig-align: center
#| out-width: '90%'
#| fig-asp: 0.6
#| echo: false

df_study1 %>%
  count(egp) %>%
  mutate(prop = (n / sum(n))*100) %>%
  ungroup() %>% 
  mutate(egp = fct_rev(egp)) %>% 
  ggplot(
    aes(x = egp, y = prop, group = egp, fill = egp)
  ) +
  geom_col(position = "stack", width = .9, alpha = 1, show.legend = F) +
  scale_y_continuous(limits = c(0, 40), 
                       breaks = seq(0, 40, 10),
                       labels = function(prop){paste0(prop, "%")},
                       expand = c(0,1)) +
  MetBrewer::scale_fill_met_d("VanGogh2") +
  coord_flip()+
  geom_text(mapping = aes(label = scales::percent(prop, accuracy = 0.1, scale = 1, decimal.mark = ",")),
              hjust = -0.2,
              fontface = "bold",
              size = 4,
              position = position_dodge(width = 0.8)) +
  labs(
    title = "Class position based on EGP scheme",
    x = NULL,
    y = NULL,
    fill = NULL,
    caption = "Source: own elaboration with data from ELSOC 2016-2023 (N obs = 7,522; N individuals = 1,317)"
  ) +
  theme_ggdist(base_size = 16)+
  theme(
    legend.position = "top",
    plot.title = element_text(face = "bold"))

```


    
- Fixed at the first wave (i.e., time-invariant) → captures average between-person differences (BE = $\bar X_{i}$)



## Estimation strategy

:::: {.incremental style="font-size: 110%;"}

**Cumulative link mixed models (CLMM)**:

- Panel structure: repeated observations (level 1) nested within individuals (level 2)

- Models both _within-person_ (WE) and _between-person_ (BE) effects; includes random effects (intercept and time slope) 

- WE/BE decomposition (person-mean centering) [@bell_fixed_2019]: 

     * Within: $X^{WE}_{it}= X_{it}-\bar{X}_i$
     * Between: $X^{BE}_{i}=\bar{X}_i$

- Formally:  

::: {.fragment}
$$\begin{aligned}
\eta_{it} = \beta_{0}+ \beta_{1}\,\text{Time}_{it} +\beta_2\,\text{Meritocracy}^{WE}_{it} +\beta_3\,\text{Meritocracy}^{BE}_{i} +\beta_4\,\mathrm{Class}^{BE}_i +
    \end{aligned}$$
    
$$\beta_5\,(\mathrm{Class}^{BE}_i\!\times\! \text{Meritocracy}^{WE}_{it}) +\beta_6\,(\mathrm{Class}^{BE}_i\!\times\! \text{Meritocracy}^{BE}_{i}) +u_{0i}+u_{1i}\,\text{Time}_{it}\\$$
:::

::::
