---
output: html_document
editor_options: 
  chunk_output_type: console
---
# Additional analyses (SEM) {background-color="#491106ff"}

```{r}
#| echo: false
#| include: false
library(here)
source(here("presentations/in_equality-Conference-2026/prod_analysis_sem.R"))
```


## Data

:::: {.incremental style="font-size: 120%;"}

- EDUMERCO: Online survey (CAWI), fielded in 2025, with adult respondents from the Metropolitan Region of Chile  

- Non-probability sample with quota-based design  
    → approximating population parameters by age, gender, education, and socioeconomic status  

- Total sample: N = 3,470  

- Analytical sample: N = 2,557 

:::: 


## Measures

:::: {.incremental style="font-size: 120%;"}

Modelo medicion

:::: 

## Analytical strategy

:::: {.incremental style="font-size: 120%;"}

- First, we estimate a Confirmatory Factor Analysis of the multidimensional meritocracy scale  
  → good model fit (CFI = 0.995, TLI = 0.991, RMSEA = 0.048)

- Second, we estimate a Structural Equation Model (MIMIC model)  
  → pension market justice regressed on latent meritocracy factors

::: {.fragment}
$$
\text{MJP}_{i} = \alpha + \beta_1 \eta^{PM}_{i} + \beta_2 \eta^{PNM}_{i} + \beta_3 \eta^{PrM}_{i} + \beta_4 \eta^{PrNM}_{i} + \gamma'X_i + \varepsilon_i
$$
:::

- Estimation: WLSMV  
  → appropriate for ordinal (Likert-type) data [@kline_principles_2023]

:::: 

# Results {background-color="#491106ff"}

## Descriptive

```{r}
#| label: fig-likermerit
#| fig-cap-location: top
#| fig-cap: "Distribution of responses on the scale of perceptions and preferences regarding meritocracy"
#| out-width: '100%'
#| echo: false
#| fig-asp: 0.6

likerplot 
```


## Bivariate
```{r}
#| label: fig-cor
#| fig-cap-location: top
#| fig-cap: "Correlation matrix between market justice and meritocracy scale"
#| out-width: '100%'
#| echo: false
#| fig-asp: 0.6


M <- db %>%
  dplyr::select(just_pension, all_of(vars_m)) %>% 
  mutate_all(~as.numeric(.)) %>% 
  psych::polychoric()

diag(M$rho) <- NA

rownames(M$rho) <- c("A. Market Justice",
                     "B. Perception Effort",
                     "C. Perception Talent",
                     "D. Perception Rich parents",
                     "E. Perception Contacts",
                     "F. Preference Effort",
                     "H. Preference Talent",
                     "H. Preference Rich parents",
                     "I. Preference Contacts")

#set Column names of the matrix
colnames(M$rho) <-c("(A)", "(B)","(C)","(D)","(E)","(F)","(G)",
                    "(H)", "(I)")

testp <- cor.mtest(M$rho, conf.level = 0.95)

#Plot the matrix using corrplot
corrplot::corrplot(M$rho,
                   method = "color",
                   addCoef.col = "black",
                   type = "upper",
                   tl.col = "black",
                   col = colorRampPalette(c("#E16462", "white", "#0D0887"))(12),
                   bg = "white",
                   na.label = "-") 
```


## Models

## Preliminary conclusions