---
bibliography: "../input/bib/jusmer.bib"
csl: "../input/bib/apa6.csl"
editor_options: 
  chunk_output_type: console
---

```{r}
#| label: set
#| echo: false
#| message: false
#| warning: false

library(knitr)
knitr::opts_chunk$set(echo = TRUE, include = TRUE, warning = FALSE, message = FALSE)

table_format <- if(is_html_output()) {
  "html"
} else if(is_latex_output()) {
  "latex"
}
table_format2 <- if(is_html_output()) {
  T
} else if(is_latex_output()) {
  F
}

options(kableExtra.html.bsTable = T)
options(knitr.kable.NA = "")
```


```{r}
#| label: packages
#| include: false

if (! require("pacman")) install.packages("pacman")

pacman::p_load(tidyverse, 
               sjmisc, 
               here, 
               summarytools)


options(scipen=999)
rm(list = ls())
```

# Data, Variables and Methods

## Data

This research draws on the Chilean Longitudinal Social Survey (ELSOC), a national panel study fielded between 2016 and 2023 to examine attitudes, emotions, and behaviors among adults aged 18 to 75 in relation to social cohesion and conflict. ELSOC uses a probabilistic, stratified, clustered, multistage sampling design covering urban centers and small towns. The sampling frame is proportionally stratified into six categories of urban population size; within these strata, households were selected through multistage procedures based on 1,067 sampled blocks.

ELSOC has been administered annually since 2016, except in 2020 when fieldwork was suspended due to the COVID-19 pandemic. For this study, we use waves 2016, 2017, 2018, 2019, 2022, and 2023. The 2021 wave is excluded because the questionnaire version omitted key variables for the analysis. By wave 6, cumulative panel attrition was approximately 40%; we therefore restrict the analysis to a balanced panel, yielding a two-level data set of *N* = 5,755 observations nested within *N* = 1,027 individuals. Additional documentation on sampling, attrition, and weighting is available from the ELSOC study team, and the data are publicly accessible through the ELSOC Dataverse.

```{r}
#| label: data
#| echo: true
#| include: false

load(file = here("input/data/proc/df_study1_long_t7_balanced.RData"))

glimpse(df_study1_long_t7_b)

# Generate analytical sample

df_study1 <- df_study1_long_t7_b %>%
  select(-muestra) %>% 
  na.omit() %>% 
  mutate(ola = case_when(ola == 1 ~ 1,
                         ola == 2 ~ 2, 
                         ola == 3 ~ 3,
                         ola == 4 ~ 4,
                         ola == 6 ~ 5,
                         ola == 7 ~ 6)) %>% 
  mutate(ola = as.factor(ola),
         ola_num = as.numeric(ola),
         ola_2=as.numeric(ola)^2)

df_study1 <- df_study1 %>%
  group_by(idencuesta) %>%             # Agrupar por el identificador del participante
  mutate(n_participaciones = n()) %>%  # Contar el número de filas (participaciones) por participante
  ungroup()

df_study1 <- df_study1 %>% filter(n_participaciones>1)

# Corregir etiquetas


df_study1$just_pension <- sjlabelled::set_label(df_study1$just_pension, 
                        label = "Pension distributive justice")

df_study1$egp <- sjlabelled::set_label(df_study1$egp, 
                        label = "Social class")

df_study1$merit_effort <- sjlabelled::set_label(df_study1$merit_effort, 
                        label = "People are rewarded for their efforts")

df_study1$merit_talent <- sjlabelled::set_label(df_study1$merit_talent, 
                        label = "People are rewarded for their intelligence")


```


## Variables

### Market justice preferences for pensions

The dependent variable in this study is preferences for market justice in the pension system. This construct is measured by a single item that captures how strongly individuals endorse conditioning access to better pensions on income. Specifically, the question is: "Is it fair that high-income individuals have better pensions than lower-income individuals?" The response options range from 1 ("Strongly disagree") to 5 ("Strongly agree"), so that higher values indicate stronger support for market-justified pension inequality.

### Social class

As for the independent variables, a class scheme was constructed based on that of Erikson, Goldthorpe and Portocarero (EGP) [@erikson_casmin_1992], collapsed into three categories: (1) Working class (V+VI+VII), which serves as the reference category, (2) Intermediate class (III+IV), and (3) Service class (I+II). Additionally, a residual category is included that considers people outside the labour force according to their statements in the survey.

We employ the EGP scheme because it captures theoretically relevant dimensions of labor market position for understanding distributive justice attitudes. Unlike occupational prestige scales or purely income-based measures, EGP distinguishes positions based on employment relations and market situations. Such situations are the degree of autonomy, job security, and contractual conditions [@barozet_measurement_2021; @barone_rise_2022] that structure both contributions to the pension system and expected returns. This distinction is crucial for our analysis since different class positions imply distinct relationships with pension schemes: service class workers typically benefit from stable employment and higher contribution rates, while working class positions face greater employment instability and lower coverage. The collapsed three-category version balances theoretical meaningfulness with statistical power, maintaining the core distinction between labor contract types (labor contracts for the working class, service relationships for the service class, and mixed positions for the intermediate class) while ensuring adequate sample sizes for longitudinal analysis.

The conversion from ISCO-08 occupational codes to EGP categories was performed using the *occupar* R package, which incorporates information on employment status (self-employed versus employee) and number of employees for self-employed workers. This approach follows the theoretical logic that class position depends not only on occupational title but also on employment relations, allowing us to differentiate, for example, between small-scale self-employed workers (petty bourgeoisie, class IVc) and employers with larger workforces (class I-II). A recognized limitation in contexts like Chile is EGP's difficulty in adequately capturing labor informality, as it was designed for formalized labor markets. However, this limitation is shared by most available occupational classification schemes and does not invalidate its analytical utility for the formal sector, which constitutes the base of the contributory pension system under normative evaluation in our dependent variable.

### Meritocracy 

Following Young’s [-@young_rise_1975] distinction, this construct is operationalised into two dimensions: effort and talent. Both dimensions are measured by the agreement with statements indicating that effort and talent are rewarded in society. Specifically, meritocratic perceptions of effort are captured by the item “In Chile, people are rewarded for their efforts” (Merit: Effort), and meritocratic perceptions of talent by the item “In Chile, people are rewarded for their intelligence and abilities” (Merit: Talent). Both items are answered on a Likert scale ranging from 1 (“Strongly disagree”) to 5 (“Strongly agree”), with higher values indicating stronger perceptions that Chilean society operates according to meritocratic principles.

@tbl-summary2 shows the main variables of the study, their response categories and their frequencies.


```{r}
#| label: tbl-summary2
#| tbl-cap: "Variable descriptive statistics for the last wave (2023)"
#| tbl-cap-location: top
#| inlcude: false
#| echo: false
#| results: asis
 

t2 <- df_study1 %>% 
  filter(ola == last(ola)) %>% 
  select(just_pension, egp, merit_effort, merit_talent) 

df<-dfSummary(t2,
               plain.ascii = FALSE,
               style = "multiline",
               tmp.img.dir = "/tmp",
               graph.magnif = 0.75,
               headings = F,  # encabezado
               varnumbers = F, # num variable
               labels.col = T, # etiquetas
               na.col = F,    # missing
               graph.col = F, # plot
               valid.col = T, # n valido
               col.widths = c(30,10,10,10))

df$Variable <- NULL # delete variable column

print(df)

```

### Controls

Based on previous research [@castillo_perceptions_2025; @lindh_public_2015; @jaime-castillo_public_2013], the statistical models include sociodemographic and attitudinal controls to account for potential compositional effects in the population. We adjust for sex (1 = male, 2 = female), age (in years), educational attainment (1 = less than university, 2 = university), and political identification (1 = left, 2 = centre, 3 = right, 4 = no identification). Following @langsaether_explaining_2022, we do not control for income, as it can be conceived as a consequence of social class —that is, a post-treatment variable— and thus risks introducing collider bias into the estimates. Descriptive statistics for the control variables can be found in the Supplementary material.

## Analytical Strategy

To examine longitudinal changes in pension inequality justification and its relationship with social class and meritocratic beliefs, we employ Cumulative Link Mixed Models (CLMM) for ordinal responses. This approach is particularly well-suited for our analytical goals for several reasons. First, it preserves the ordinal nature of our dependent variable without imposing assumptions about equal spacing between response categories. Second, it accounts for the hierarchical structure of our panel data, with repeated observations (Level 1) nested within individuals (Level 2). Third, it enables the decomposition of variance into within-person (WE) and between-person (BE) components, allowing us to distinguish temporal changes within individuals from stable differences between individuals [@singer_applied_2003; @bell_fixed_2019].

### Within-Person and Between-Person Random Effects Models

Following the recommendations of @bell_explaining_2015, we decompose time-varying predictors into their within-person and between-person components using person-mean centering:

$$
X_{WE_{it}} = X_{it} - \bar{X}_i
$$

$$
X_{BE_i} = \bar{X}_i
$$

where $X_{it}$ represents the observed value for individual $i$ at time $t$, $\bar{X}_i$ is the person-specific mean across all observations, $X_{WE_{it}}$ captures within-person deviations from the individual mean (representing temporal fluctuations), and $X_{BE_i}$ represents stable between-person differences (time-invariant characteristics).

This decomposition allows us to distinguish two theoretically distinct processes: (1) Within-person effects capture how changes in an individual's meritocratic beliefs over time relate to changes in their pension inequality justification, controlling for stable individual characteristics. (2) Between-person effects capture how individuals with different average levels of meritocratic beliefs differ in their pension inequality justification, net of temporal fluctuations.

### Model Specification

The general form of our cumulative link mixed model is:

$$
\begin{split}
\eta_{it} = \beta_0 + \beta_1 \text{time}_{it} + \beta_2 X_{WE_{it}} + \beta_3 X_{BE_i} + \beta_4 \text{Class}_i + \\
\beta_5 (\text{Class}_i \times X_{WE_{it}}) + \beta_6 (\text{Class}_i \times X_{BE_i}) + u_{0i} + u_{1i} \text{time}_{it}
\end{split}
$$

where $\eta_{it}$ is the linear predictor for individual $i$ at time $t$, linked to the cumulative probabilities through a logit link function; $\beta_0$ is the fixed intercept (threshold); $\beta_1$ captures the linear time trend; $\beta_2$ and $\beta_3$ represent within-person and between-person effects of meritocratic perceptions, respectively; $\beta_4$ captures social class between effects; $\beta_5$ and $\beta_6$ represent interactions between social class and within/between-person meritocratic perceptions; $u_{0i}$ is the random intercept (capturing unobserved individual heterogeneity); and $u_{1i}$ is a random slope for time (allowing individual-specific temporal trajectories).

The random effects are assumed to follow a bivariate normal distribution:

$$
\begin{pmatrix} u_{0i} \\ u_{1i} \end{pmatrix} \sim N \begin{pmatrix} \begin{pmatrix} 0 \\ 0 \end{pmatrix}, \begin{pmatrix} \sigma^2_{u_0} & \sigma_{u_{01}} \\ \sigma_{u_{01}} & \sigma^2_{u_1} \end{pmatrix} \end{pmatrix}
$$

### Model Building Strategy

We estimate a series of nested models to test our hypotheses:

1. **Model 1 (Time Effects)**: Baseline model examining temporal trends with linear and quadratic time terms
2. **Model 2 (Class Effects)**: Adds social class as a predictor
3. **Model 3 (Within-Person Effects)**: Includes within-person meritocratic beliefs
4. **Model 4 (Between-Person Effects)**: Adds between-person meritocratic beliefs
5. **Model 5 (Full Model with Controls)**: Incorporates sociodemographic controls (education, political identification, sex, age)
6. **Model 6 (Interactions)**: Tests interactions between social class and meritocratic beliefs (both WE and BE)

All models are estimated using maximum likelihood with the Hessian matrix computed explicitly to obtain standard errors. Longitudinal survey weights are applied to account for the complex sampling design and panel attrition. Models are implemented using the `clmm()` function from the `ordinal` package in R (@haubo_cumulative_2018).



