```{r}
#| echo: false
#| message: false
#| warning: false

library(knitr)
knitr::opts_chunk$set(echo = FALSE, include = TRUE, warning = FALSE, message = FALSE)

table_format <- if(is_html_output()) {
  "html"
} else if(is_latex_output()) {
  "latex"
}
table_format2 <- if(is_html_output()) {
  T
} else if(is_latex_output()) {
  F
}

options(kableExtra.html.bsTable = T)
options(knitr.kable.NA = "")
```

```{r}
#| include: false

if (! require("pacman")) install.packages("pacman")

pacman::p_load(tidyverse, 
               sjmisc, 
               sjPlot, 
               here, 
               srvyr,
               ggdist,
               kableExtra,
               ggalluvial, 
               MetBrewer,
               patchwork,
               sjlabelled)


options(scipen=999)
rm(list = ls())
```

```{r}
#| echo: false

load(file = here("input/data/proc/df_balanced.RData"))

# Generate analytical sample

df_study1 <- df_balanced %>%
  select(-muestra) %>% 
  na.omit() %>% 
  mutate(ola = case_when(ola == 1 ~ 1,
                         ola == 2 ~ 2, 
                         ola == 3 ~ 3,
                         ola == 4 ~ 4,
                         ola == 6 ~ 5,
                         ola == 7 ~ 6)) %>% 
  mutate(ola = as.factor(ola),
         ola_num = as.numeric(ola),
         ola_2=as.numeric(ola)^2)

df_study1 <- df_study1 %>%
  group_by(idencuesta) %>%             # Agrupar por el identificador del participante
  mutate(n_participaciones = n()) %>%  # Contar el número de filas (participaciones) por participante
  ungroup()

df_study1 <- df_study1 %>% filter(n_participaciones>1)

# Corregir etiquetas

df_study1$just_pension <- sjlabelled::set_label(df_study1$just_pension, 
                        label = "Pension distributive justice")

df_study1$egp <- sjlabelled::set_label(df_study1$egp, 
                        label = "Social class")

df_study1$merit_effort <- sjlabelled::set_label(df_study1$merit_effort, 
                        label = "People are rewarded for their efforts")

df_study1$merit_talent <- sjlabelled::set_label(df_study1$merit_talent, 
                        label = "People are rewarded for their intelligence")

```

## Descriptive analysis

@fig-alluvial depicts the temporal evolution of market justice preferences in pensions from 2016 to 2023. Each year presents percentage frequencies, and the flows between them reflect changes in opinion among the same people from one year to the next. The results indicate substantial attitudinal fluidity, with extensive cross-flows suggesting that many respondents revise their views over time rather than holding fixed positions. Most changes occur between adjacent categories, consistent with gradual shifts, although some movements span multiple categories. For example, among those who strongly disagreed with income-based pension access in 2019, about 15.2% remained in that position by 2022, while the remaining 12.5% shifted to other categories, primarily to disagreement or agreement. Overall, the majority continue to disagree with the notion that it is fair for higher-income people to receive better pensions than those with lower incomes. Yet, agreement has increased over time: the combined share of agree + strongly agree rises from 16.9% in 2019 to 28.4% in 2023 (an increase of roughly 11.5 percentage points). Transitions appear particularly turbulent during the 2019–2022 period—overlapping with social unrest and the pandemic—suggesting that these attitudes are responsive to major contextual shocks and shifts in pension-related public discourse.


```{r}
#| label: fig-alluvial
#| out-width: '100%'
#| fig-asp: 1
#| fig-cap: "Changes in the justification of inequality in pensions over time (2016-2023)"
#| fig-cap-location: bottom
#| echo: false


datos.pension <- df_study1 %>% 
   mutate(just_pension = factor(just_pension, 
                              levels = c("Strongly agree",
                                         "Agree",
                                         "Neither agree nor disagree",
                                         "Disagree",
                                         "Strongly disagree"))) %>% 
  group_by(idencuesta, ola) %>% 
  count(just_pension) %>% 
  group_by(ola) %>% 
  mutate(porcentaje=n/sum(n)) %>% 
  ungroup() %>% 
  na.omit() %>% 
  mutate(wave = case_when(ola == 1 ~ "2016",
                          ola == 2 ~ "2017",
                          ola == 3 ~ "2018",
                          ola == 4 ~ "2019",
                          ola == 5 ~ "2022",
                          ola == 6 ~ "2023"),
         wave = factor(wave, levels = c("2016",
                                        "2017",
                                        "2018",
                                        "2019",
                                        "2022",
                                        "2023")))



etiquetas.pension <- df_study1 %>%
  mutate(just_pension = factor(just_pension, 
                              levels = c("Strongly agree",
                                         "Agree",
                                         "Neither agree nor disagree",
                                         "Disagree",
                                         "Strongly disagree"))) %>% 
  group_by(ola, just_pension) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(ola) %>%
  mutate(porcentaje = count / sum(count)) %>% 
  na.omit() %>% 
  mutate(idencuesta = 1,
         wave = case_when(ola == 1 ~ "2016",
                          ola == 2 ~ "2017",
                          ola == 3 ~ "2018",
                          ola == 4 ~ "2019",
                          ola == 5 ~ "2022",
                          ola == 6 ~ "2023"),
         wave = factor(wave, levels = c("2016",
                                        "2017",
                                        "2018",
                                        "2019",
                                        "2022",
                                        "2023")))

datos.pension %>% 
  ggplot(aes(x = wave, fill = just_pension, stratum = just_pension,
             alluvium = idencuesta, y = porcentaje)) +
  ggalluvial::geom_flow(alpha = .4) + 
  ggalluvial::geom_stratum(linetype = 0) +
  scale_y_continuous(labels = scales::percent) + 
  scale_fill_manual(values =  c("#0571B0","#92C5DE","#b3b3b3ff","#F4A582","#CA0020")) +
  geom_text(data = etiquetas.pension,
            aes(label = ifelse(porcentaje > 0 , scales::percent(porcentaje, accuracy = .1),"")),
            position = position_stack(vjust = .5),
            show.legend = FALSE,
            size = 3,
            color = 'white')+
  labs(y = "%",
       x = NULL,
       fill = NULL,
       title = NULL) +
  theme_ggdist() +
  theme(legend.position = "bottom") +
  labs(caption = "Source: Elaborated based on ELSOC (2016–2023)")

```


## Bivariate relationships

@fig-scatter-corr shows the correlation matrix between meritocratic perceptions and preferences for pension market justice. Both effort-based ($r$ = 0.10, $p$ < 0.001) and talent-based ($r$ = 0.08, $p$ < 0.001) meritocratic perceptions are positively associated with greater support for market-based pension welfare. Although these correlations are modest, their statistical significance indicates a consistent bivariate relationship. The two meritocracy indicators are also strongly correlated ($r$ = 0.68, $p$ < 0.001), suggesting they capture a shared meritocratic worldview while remaining sufficiently distinct to justify treating them as separate constructs. Taken together, these patterns provide initial support for the expectation that meritocratic perceptions are associated with greater acceptance of income-linked pension outcomes, while also suggesting that other factors likely account for most of the variation in these preferences.

```{r}
#| label: fig-scatter-corr
#| fig-width: 8
#| fig-height: 6
#| fig-cap: "Correlation matrix between meritocratic beliefs and market justice"
#| fig-cap-location: top
#| echo: false

# Convertir variables a numéricas
df_scatter <- df_study1 %>%
  mutate(
    just_pension_num = as.numeric(just_pension),
    merit_effort_num = as.numeric(merit_effort),
    merit_talent_num = as.numeric(merit_talent)
  )

# Calcular correlaciones
cor_data <- df_scatter %>%
  select(just_pension_num, merit_effort_num, merit_talent_num) %>%
  na.omit()

# Correlaciones de Pearson
cor_matrix <- cor(cor_data, method = "pearson")

# Crear visualización con ggplot
cor_df <- as.data.frame(cor_matrix) %>%
  rownames_to_column("var1") %>%
  pivot_longer(-var1, names_to = "var2", values_to = "correlation")

# Calcular p-valores
n <- nrow(cor_data)
cor_df <- cor_df %>%
  mutate(
    t_stat = correlation * sqrt(n - 2) / sqrt(1 - correlation^2),
    pvalue = 2 * pt(abs(t_stat), df = n - 2, lower.tail = FALSE),
    sig = case_when(
      pvalue < 0.001 ~ "***",
      pvalue < 0.01 ~ "**",
      pvalue < 0.05 ~ "*",
      TRUE ~ ""
    ),
    label = paste0(sprintf("%.2f", correlation), sig),
    var1 = factor(var1, levels = c("just_pension_num", "merit_effort_num", "merit_talent_num"),
                  labels = c("Market Justice\nin Pensions", "Merit:\nEffort", "Merit:\nTalent")),
    var2 = factor(var2, levels = c("just_pension_num", "merit_effort_num", "merit_talent_num"),
                  labels = c("Market Justice\nin Pensions", "Merit:\nEffort", "Merit:\nTalent"))
  )

ggplot(cor_df, aes(x = var2, y = var1, fill = correlation)) +
  geom_tile(color = "white", linewidth = 1) +
  geom_text(aes(label = label), size = 6, fontface = "bold") +
  scale_fill_gradient2(low = "#0571B0", mid = "white", high = "#CA0020",
                       midpoint = 0, limits = c(-1, 1),
                       name = "Pearson r") +
  labs(x = NULL, y = NULL,
       caption = "*** p < 0.001; ** p < 0.01; * p < 0.05\nSource: Elaborated based on ELSOC (2016–2023)") +
  theme_minimal() +
  theme(text = element_text(size = 12),
        axis.text.x = element_text(angle = 0, hjust = 0.5, size = 11),
        axis.text.y = element_text(size = 11),
        legend.position = "right",
        plot.caption = element_text(size = 10, face = "italic", hjust = 0))
```

Turning to social class differences, @fig-meanchange plots longitudinal trajectories of pension market justice by class from 2016 to 2023. The figure shows a broadly upward trend across all classes, especially after 2019, alongside increasingly similar levels over time. The unemployed group shows the most significant change, moving from strong rejection (approximately −0.6 in 2016) to moderate acceptance (approximately 0.7 in 2023). Retirees/pensioners also increase steadily, from near zero in 2016 to roughly 0.5 by 2023, and the service class rises notably—particularly between 2019 and 2022—reaching values close to 0.5. The working class follows a similar trajectory, though at slightly lower levels. Overall, the trajectories suggest convergence toward greater acceptance of pension inequality during 2019–2023, a period marked by social unrest, the pandemic, and pension fund withdrawals (*10% retiros*), raising questions about how these disruptions and the surrounding policy debate reshaped views of distributive justice.

```{r}
#| label: fig-meanchange
#| out-width: '100%'
#| fig-asp: 0.5
#| fig-cap: "Changes in the standarized mean of market justice preferences in pensions by social class (2016-2023)"
#| fig-cap-location: top
#| echo: false
#| results: asis

df_pond <- df_study1 %>% 
  mutate(ola = case_when(ola == 1 ~ "2016",
                         ola == 2 ~ "2017",
                         ola == 3 ~ "2018",
                         ola == 4 ~ "2019",
                         ola == 5 ~ "2022",
                         ola == 6 ~ "2023"),
         ola = factor(ola, levels = c("2016",
                                      "2017",
                                      "2018",
                                      "2019",
                                      "2022",
                                      "2023"))) %>% 
  as_survey_design(.data = .,
                   ids = segmento, 
                   strata = estrato, 
                   weights = ponderador_long_total)

df_pond <- df_pond %>% 
  select(ola, egp, just_pension) %>% 
  mutate(just_pension = as.numeric(just_pension)) %>% 
  mutate(just_pension_z = as.numeric(scale(just_pension)))


shapes_egp <- c(
  "Working class (V+VI+VII)"      = 16,  # circle
  "Intermediate class (III+IV)" = 17,  # triangle
  "Service class (I+II)"      = 15,  # square
  "Retired or pensioner"            = 18,  # diamond
  "Unemployed"         = 8,   # star
  "Performs unpaid tasks"        = 3    # plus
)


base_cols <- c("#1F77B4","#6BAED6","#08306B","#BDBDBD","#D7301F","#FC8D59")
cols_egp <- setNames(base_cols, levels(df_pond$egp))

df_pond %>% 
  select(ola, egp, just_pension_z) %>% 
  group_by(ola, egp) %>% 
  summarise_all(~survey_mean(., vartype = "ci")) %>% 
  ggplot(aes(x = ola, y = just_pension_z, group = egp)) +
  geom_point(aes(shape = egp, color = egp), size = 3.5) +
  geom_line(aes(color = egp), linewidth = 0.8) +
  geom_errorbar(aes(ymin = just_pension_z_low, ymax = just_pension_z_upp, color = egp),
                width = 0.1) +
  scale_shape_manual(values = shapes_egp) +
  scale_color_manual(values = cols_egp) +
  scale_y_continuous(limits = c(-1, 1), n.breaks = 5) +
  labs(y = "Mean (Z-score)", x = "Wave", color = NULL, shape = NULL,
       caption = "Source: own elaboration with pooled data from ELSOC 2016-2023 (N obs = 5,755; N groups = 1,027)") +
  theme_ggdist() +
  theme(legend.position = "top", text = element_text(size = 12))

 
```


